---
title: "R coding Exercise"
output: 
  html_document:
    toc: FALSE
    theme: readable
    highlight: tango
---

#View Data
```{r}

#Load "dslabs" package to manipulate data
library("dslabs")
library("tidyverse")
library("broom")

#Check out the help file in gapminder
help(gapminder)

#Lets get an overview of the data
str(gapminder)

#Now for a summary of the data
summary(gapminder)

#To see what type of object gapminder is, we use the class function
class(gapminder)

```
#Creating a data subset

```{r}

#Now let look at only the African countries
africadata <- gapminder[ which(gapminder$continent=="Africa"), ]

#Lets get an overview & summary of the new data
str(africadata)
summary(africadata)

#Making two new data frames
imle <- africadata[c("infant_mortality","life_expectancy")]
ple <- africadata[c("population","life_expectancy")]

#Overview of new sets
str(imle)
str(ple)

#Summary of new sets
summary(imle)
summary(ple)

```

#Plotting the variables

```{r}
#life expectancy vs. infant mortality
ggplot(imle, aes(infant_mortality, life_expectancy)) + geom_point()

#Here we have a negative correlation

#life expectancy vs. population
ggplot(ple, aes(population, life_expectancy)) + geom_point() +scale_x_log10()

#Here we have a loose positive correlation. Notice the "streaks" in the data
```

#Data cleaning
```{r}

#Use the is.na function to see what data we are missing
africadata$year[is.na(africadata$infant_mortality)]

#Creating a datafram that only includes data from year 2000
newyr <- africadata[africadata$year=="2000", c("country", "year", "infant_mortality", "life_expectancy", "fertility", "population", "gdp", "continent", "region")]

#Check the structure and summary
str(newyr)
summary(newyr)
```
#Plotting with new subset
```{r}

#Plot new dataframe, same variable as before
#Life expectancy vs. Infant Mortality 2000 plot
plot(infant_mortality ~ life_expectancy, data=newyr)

#Life expectancy vs. Population 2000 plot
plot(life_expectancy ~ population, data=newyr, log = "x")
#no relationship

#Lets create fitted lines for the life expectancy outcome

#Life expectancy vs. Infant Mortality 2000
fit1<- lm(formula = life_expectancy ~ infant_mortality, data = newyr)

#Life expectancy vs. Population 2000
fit2<- lm(formula = life_expectancy ~ population, data = newyr)

#summary(fit1)

#summary(fit2)

```
#After looking at P-values for both, it appears that the correlation between Infant Mortality and Life Expectancy has a P-val <0.05 and is a better predictor than Population, which has a P-val >0.05.

#We've established that infant mortality is a good predictor of life expectancy. Other variables in this dataset include fertility and GDP - how do these variables relate to life expectancy? Could they be better predictors of life expectancy than infant mortality is? Let's explore this. 

#Subset data
```{r}
#create subset of africadata containing fertility and life expectancy
fle <- africadata %>% select("fertility", "life_expectancy")

#create subset of africadata containing GDP and life expectancy
gdple <- africadata %>% select("gdp", "life_expectancy")
```

#Create plots
```{r}
#create plot of fertility vs. life expectancy
ggplot(fle, aes(fertility, life_expectancy)) + geom_point()

#Here, we have a weak negative correlation

#create plot of GDP vs. life expectancy
ggplot(gdple, aes(gdp, life_expectancy)) + geom_point()

#Here, we see a possible weak positive correlation, with a large of cluster of values on the lower end of the GDP range

```
#Fit a model
```{r}
#Let's look at how fertility and GDP perform as predictors of life expectancy using a linear model

#Life expectancy vs fertility
fit3 <- lm(formula = life_expectancy ~ fertility, data = fle)

#Life expectancy vs GDP
fit4 <- lm(formula = life_expectancy ~ gdp, data = gdple)

#view summaries of fit3 and fit4
summary(fit3)
summary(fit4)

```
The summaries show that both fit3 and fit4 have a p-value of < 0.05, indicating that fertility and GDP may be good predictors of life expectancy. Let's display the `lm()` output as a table using the `broom` package.

#Use `broom` to view fit results
```{r}
#View fertility vs life expectancy (fit3) results
tidy(fit3)

#View GDP vs life expectancy (fit4) results
tidy(fit4)

```
After exploring the relationship between fertility and life expectancy and GDP and life expectancy, it appears that both fitted models have p-values of <0.05, indicating that fertility and GDP may also have an informative/predictive relationship with life expectancy. 
